---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# DR4PL

The package DR4PL (Dose Response 4 Parameter Logisitic model) specializes in applying the 4PL model. The goal of DR4PL is to bring a statistical method which is capable of handeling specific error cases of which other statistical packages produce errors. The package DR4PL also alows for the user to decide how their theta variable is approximated. The user may choose the default logistic model or use Mead's Method. Additionally, the user may decide between four loss functions to minimize: Squared, Absolute, Huber, Tukey's biweight.

## Installation

You can install DR4PL from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ahwbest/dr4pl")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```
